# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

```{r}
unzip("activity.zip")
activity <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?

```{r}
library(reshape)
steps <- tapply(activity$steps,activity$date,sum,na.rm=TRUE)
steps<- melt(steps)
colnames(steps) <- c("Date","steps")
p1<- hist(steps$steps,breaks=10,xlab="Steps per day", ylab="Count",
     main="Total number of steps taken per day")
```

```{r}
summary(steps$steps)
```

## What is the average daily activity pattern?

```{r}
library(ggplot2)
dpattern <- aggregate(steps ~ interval, data=activity,mean, na.rm=TRUE)
colnames(dpattern) <- c("interval","avg.steps")
qplot(x=dpattern$interval,y=dpattern$avg.steps,geom=c("line"),
      xlab="Average number of steps",ylab="5 minute interval",
      main="Average daily pattern")
```
```{r}
dpattern[which.max(dpattern$avg.steps),]
```
## Imputing missing values

```{r}
missingvals <- is.na(activity$steps)
```
The total missing values in the activity data set is `r sum(missingvals)`

```{r}
fillvals <- function(steps,interval) {
  if(!is.na(steps))
  {
    return(steps)
  } else {
    return(dpattern[dpattern$interval == interval,"avg.steps"])
  }
}
activity$steps.v2 <- mapply(fillvals,c(activity$steps),activity$interval)
```

```{r}
steps <- tapply(activity$steps.v2,activity$date,sum,na.rm=TRUE)
steps<- melt(steps)
colnames(steps) <- c("Date","steps")
p2<- hist(steps$steps,breaks=10,xlab="Steps per day", ylab="Count",
     main="Total number of steps taken per day (filled)")
```

```{r}
summary(steps$steps)
```

Mean and median are higher. If we compare with the previous histogram, we notice that we are removing all the zero values given by the missing values, and replacing by the mean of the given interval.

```{r}
plot(p2,col=rgb(0,0,1,1/4))
plot(p1,col=rgb(1,0,0,1/4),add=TRUE)
```
The main effect is in the bins arround the zero values and the the bins arround the mean value.

## Are there differences in activity patterns between weekdays and weekends?

First we create the new factor variable week with two leves: weekend and weekday

```{r}
weekdays.factors <- weekdays(as.Date(activity$date))
activity$week <- factor(by(weekdays.factors,1:length(weekdays.factors),
                function(row){
                  if( row %in% c("Saturday","Sunday"))
                  {
                    return(c("weekend"))
                  } else {
                    return(c("weekday"))
                  }
                  })
                )

```

```{r}
dpattern <- aggregate(steps ~ interval+week, data=activity,mean, na.rm=TRUE)
qplot(interval,steps,data=dpattern, geom=c("line"),facets=week~.,
      xlab="5 minute interval",ylab="Number of steps")
```